Clustering in a hyperbolic model of complex networks.

### 2020-03-11 hyperbolic_degrees.tex
small corrections

Now let $y > 2 \log(\pi/2)$ and recall that $\mu(y) = \xi e^{\frac{y}{2}}$ where $\xi = \frac{4\alpha \nu}{(2\alpha - 1)\pi}$. Then, after some simple algebra, we have that
\begin{align*}
	\mu_{\mathrm{box}}(y)
	&= \int_0^{h(y)} \int_{-\frac{\pi}{2}e^{\frac{R}{2}}}^{\frac{\pi}{2}e^{\frac{R}{2}}} 
		\ind{|x^\prime| \le e^{\frac{y+y^\prime}{2}}} f(x^\prime,y^\prime) \, dx^\prime \, dy^\prime
		+ \int_{h(y)}^{R} \int_{-\frac{\pi}{2}e^{\frac{R}{2}}}^{\frac{\pi}{2}e^{\frac{R}{2}}} 
		f(x^\prime,y^\prime) \, dx^\prime \, dy^\prime\\
	&= \frac{2 \alpha \nu}{\pi} e^{\frac{y}{2}} \int_0^{h(y)} e^{-(\alpha - \frac{1}{2})y^\prime} \, dy^\prime
		+ \alpha \nu e^{\frac{R}{2}} \int_{h(y)}^{R} e^{-\alpha y^\prime} \, dy^\prime \\
	&= \xi e^{\frac{y}{2}}\left(1 - \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} 
		e^{-(\alpha - \frac{1}{2})(R - y)}\right)
	+ \nu e^{\frac{R}{2}}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} e^{-\alpha(R - y)} 
		- e^{-\alpha R}\right)\\
	&= \mu(y)\left(1 - \phi_n(y) \right).
\end{align*}
where 
\[
	\phi_n(y) :=  \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} \hspace{-3pt} e^{-(\alpha - \frac{1}{2})(R - y)}
				+ \frac{\nu}{\xi}e^{-(\alpha - \frac{1}{2})R - \frac{y}{2}} - \frac{\nu}{\xi}\left(\frac{\pi}{2}\right)^{-2\alpha} e^{-(\alpha-\frac{1}{2})(R - y)}.
\]
Since $R - y \ge \varepsilon R$ we have that $|\phi_n(y)|$ is uniformly bounded by
$\bigO{e^{-(\alpha - \frac{1}{2})\varepsilon R}}$, which is $\smallO{1}$ for $\alpha > \frac{1}{2}$. 
\end{proof}

@@

\begin{proof}[Proof of Lemma~\ref{lem:factmoment}]
We give the proof for the Poissonized KPKVB model. The proof for the finite box model $\Gbox$ follows using similar arguments.

@@

Using Lemma~\ref{lem:asympind}, we conclude that
\begin{align*}
	\Exp{ { N_{\Po}(k_n) \choose r}} 
	&= \frac{1}{r!} \int_\Rcal \cdots \int_\Rcal 
		\varphi_{\Po}(\{p_1,\dots,p_r\}, k_n; p_1,\dots,p_r) 
		\dd \mu(p_1) \cdots \dd \mu(p_{r}) \\
	&= (1+o(1))\frac{1}{r!}\left(\int_\Rcal \varphi_{\Po}(\{p_1\}, k_n;p_1) f(x_1,y_1) \dd x_1 \dd y_1 \right)^r \\
	&= (1+o(1))\frac{1}{r!}(\Exp{N_{\Po}(k_n)})^r.
\end{align*}
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{thm:degrees_hyperbolic}]

@@

For $k_n \ge n^{1 - \varepsilon}$ we have, by our choice of $\varepsilon$, that $\frac{3}{2} - (2\alpha + 1)(1-\varepsilon) < 0$, and thus
\[
	\sqrt{\frac{n}{n-k_n}} n k_n^{-(2\alpha + 1)} = \bigO{n^{\frac{3}{2}} k_n^{-(2\alpha + 1)}}
	= \bigO{n^{\frac{3}{2} - (2\alpha + 1)(1-\varepsilon)}} = \smallO{1}.
\]
\end{enumerate}
\end{proof}

>>>>

Now let $y > 2 \log(\pi/2)$ and recall that $\mu(y) = \xi e^{\frac{y}{2}}$ where $\xi = \frac{4\alpha \nu}{(2\alpha - 1)\pi}$. Then, after some simple algebra, we have that
\begin{align*}
	\mu_{\mathrm{box}}(y)
	&= \int_0^{h(y)} \int_{-\frac{\pi}{2}e^{\frac{R}{2}}}^{\frac{\pi}{2}e^{\frac{R}{2}}} 
		\ind{|x^\prime| \le e^{\frac{y+y^\prime}{2}}} f(x^\prime,y^\prime) \, dx^\prime \, dy^\prime
		+ \int_{h(y)}^{R} \int_{-\frac{\pi}{2}e^{\frac{R}{2}}}^{\frac{\pi}{2}e^{\frac{R}{2}}} 
		f(x^\prime,y^\prime) \, dx^\prime \, dy^\prime\\
	&= \frac{2 \alpha \nu}{\pi} e^{\frac{y}{2}} \int_0^{h(y)} e^{-(\alpha - \frac{1}{2})y^\prime} \, dy^\prime
		+ \alpha \nu e^{\frac{R}{2}} \int_{h(y)}^{R} e^{-\alpha y^\prime} \, dy^\prime \\
	&= \xi e^{\frac{y}{2}}\left(1 - \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} 
		e^{-(\alpha - \frac{1}{2})(R - y)}\right)
	+ \nu e^{\frac{R}{2}}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} e^{-\alpha(R - y)} 
		- e^{-\alpha R}\right)\\
	&= \mu(y)\left(1 - \phi_n(y) \right).
\end{align*}
where 
\begin{equation}\label{eq:def_average_degree_Gbox_phi}
	\phi_n(y) :=  \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)} \hspace{-3pt} e^{-(\alpha - \frac{1}{2})(R - y)}
				+ \frac{\nu}{\xi}e^{-(\alpha - \frac{1}{2})R - \frac{y}{2}} - \frac{\nu}{\xi}\left(\frac{\pi}{2}\right)^{-2\alpha} e^{-(\alpha-\frac{1}{2})(R - y)}.
\end{equation}
Since $R - y \ge \varepsilon R$ we have that $|\phi_n(y)|$ is uniformly bounded by
$\bigO{e^{-(\alpha - \frac{1}{2})\varepsilon R}}$, which is $\smallO{1}$ for $\alpha > \frac{1}{2}$. 
\end{proof}

@@

\begin{proofof}{Lemma~\ref{lem:factmoment}}
We give the proof for the Poissonized KPKVB model. The proof for the finite box model $\Gbox$ follows using similar arguments.

@@

Using Lemma~\ref{lem:asympind}, we conclude that
\begin{align*}
	\Exp{ { N_{\Po}(k_n) \choose r}} 
	&= \frac{1}{r!} \int_\Rcal \cdots \int_\Rcal 
		\varphi_{\Po}(\{p_1,\dots,p_r\}, k_n; p_1,\dots,p_r) 
		\dd \mu(p_1) \cdots \dd \mu(p_{r}) \\
	&= (1+o(1))\frac{1}{r!}\left(\int_\Rcal \varphi_{\Po}(\{p_1\}, k_n;p_1) f(x_1,y_1) \dd x_1 \dd y_1 \right)^r \\
	&= (1+o(1))\frac{1}{r!}(\Exp{N_{\Po}(k_n)})^r.
\end{align*}
\end{proofof}

@@

\begin{proofof}{Theorem~\ref{thm:degrees_hyperbolic}}

@@

For $k_n \ge n^{1 - \varepsilon}$ we have, by our choice of $\varepsilon$, that $\frac{3}{2} - (2\alpha + 1)(1-\varepsilon) < 0$, and thus
\[
	\sqrt{\frac{n}{n-k_n}} n k_n^{-(2\alpha + 1)} = \bigO{n^{\frac{3}{2}} k_n^{-(2\alpha + 1)}}
	= \bigO{n^{\frac{3}{2} - (2\alpha + 1)(1-\varepsilon)}} = \smallO{1}.
\]
\end{enumerate}
\end{proofof}

### 2020-03-11 preliminaries.tex
small corrections

With this notation we then have that $\BallPo{p} \cap \Pcal$ denotes the set of neighbours of a vertex $p \in \Ginf$.
We will denote the intensity measure of the Poisson process $\mathcal{P}$ by $\mu = \mu_{\alpha, \nu}$, i.e. for every 
Borel-measurable subset $S \subseteq \eR^2$ we have $\mu(S) = \int_S f(x,y) \dd x \dd y$.

@@

In particular, if $\Po(\lambda)$ denotes a Poisson random variable with mean $\lambda$, we have the following Chernoff bound (c.f. \cite[Lemma 1.2]{penrose2003random})
\begin{equation}\label{eq:def_chernoff_bound_poisson}
	\Prob{\left|\mathrm{Po}(\lambda) - \lambda\right| \ge x} \le 2e^{-\frac{x^2}{2(\lambda + x)}}.
\end{equation}
In particular, if $\lambda = \lambda_n \to \infty$, then for any $C>0$,
\begin{equation}\label{eq:def_chernoff_bound_poisson_C}
	\Prob{\left|\mathrm{Po}(\lambda_n) - \lambda_n\right| \ge C \sqrt{\lambda_n\log(\lambda_n)}} \le \bigO{\lambda_n^{-\frac{C^2}{2}}}.
\end{equation}

@@

\begin{proposition}[Concentration of heights]\label{prop:concentration_height_general}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n}$. Furthermore, let $\hat{\mu}(y)$ denoting either $\Mu{\BallHyp{y}}$, $\Mu{\BallPon{y}}$ or $\Mu{\BallPo{y}}$. Then for any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, it holds that
\begin{equation}\label{eq:error_bound_int_rho_not_K}
	\int_0^\infty \hspace{-3pt} h(y) \Prob{\Po(\hat{\mu}(y)) = k_n} \alpha e^{-\alpha y} \dd y
	\sim \int_{\Kcal_C(k_n)} \hspace{-8pt} h(y) \Prob{\Po(\hat{\mu}(y)) = k_n} \alpha e^{-\alpha y} \dd y,
\end{equation}
as $n \to \infty$.
\end{proposition}

>>>>

With this notation we then have that $\BallPo{p} \cap \Pcal$ denotes the set of neighbours of a vertex $p \in \Ginf$.
We will denote the intensity measure of the Poisson process $\mathcal{P}$ by $\mu = \mu_{\alpha, \nu}$, i.e. for every 
Borel-measurable subset $S \subseteq \eR^2$ we have $\mu(S) = \int_S f(x,y) \dd x \dd y$. Using the notation $p = (x,y)$ for a point in $\R \times \R_+$ we shall write $\int_S h(p) \dd \mu(p)$ for the integral of $h$ over $S$ with respect to the intensity measure $\mu$, i.e. $\int_S h(p) \dd \mu(p) = \int_{S} h(x,y) f(x,y) \dd x \dd y$.

@@

Let $\Po(\lambda)$ denote a Poisson random variable with mean $\lambda$. Then we have the following Chernoff bound (c.f. \cite[Lemma 1.2]{penrose2003random})
\begin{equation}\label{eq:def_chernoff_bound_poisson}
	\Prob{\left|\mathrm{Po}(\lambda) - \lambda\right| \ge x} \le 2e^{-\frac{x^2}{2(\lambda + x)}}.
\end{equation}
In particular, if $\lambda = \lambda_n \to \infty$, then for any $C>0$,
\begin{equation}\label{eq:def_chernoff_bound_poisson_C}
	\Prob{\left|\mathrm{Po}(\lambda_n) - \lambda_n\right| \ge C \sqrt{\lambda_n\log(\lambda_n)}} \le \bigO{\lambda_n^{-\frac{C^2}{2}}}.
\end{equation}

@@

\begin{proposition}[Concentration of heights]\label{prop:concentration_height_general}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n}$. Furthermore, let $\hat{\mu}(y)$ denoting either $\Mu{\BallHyp{y}}$, $\Mu{\BallPon{y}}$ or $\Mu{\BallPo{y}}$. Then for any continuous function $h : \R_+ \rightarrow  \R$, such that $h(y) = \bigO{e^{\beta y}}$ as $y \to \infty$ for some $\beta < \alpha$, it holds that
\begin{equation*}%\label{eq:error_bound_int_rho_not_K}
	\int_0^\infty \hspace{-3pt} h(y) \Prob{\Po(\hat{\mu}(y)) = k_n} \alpha e^{-\alpha y} \dd y
	\sim \int_{\Kcal_C(k_n)} \hspace{-8pt} h(y) \Prob{\Po(\hat{\mu}(y)) = k_n} \alpha e^{-\alpha y} \dd y,
\end{equation*}
as $n \to \infty$.
\end{proposition}

### 2020-03-11 proofs_infinite_k.tex
small corrections

Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:mainktoinfty} based on the different propositions described below, plus the sections in which theses propositions are proved. Observe that the order in which the intermediate results are proved is reversed with respect to the natural order of reasoning. This does not create any circular logic, since each intermediate result is independent of the others. We choose this order because results proved in the later stages are helpful to deal with error terms coming up in proofs at earlier stages and hence help streamline those proofs. Below briefly describe each of the intermediate steps leading up to the proof of Theorem~\ref{thm:mainktoinfty}. We start with an observation on the dependence between the degree of a point $p = (x,y)$ and its height $y$.
%=======
%Notice that the only difference between $c(k;G)$ and $c^\ast(k;G)$ is that we replace $N_k$ by its expectation $\Exp{N_k}$. The advantage is that now, the only randomness is in the formation of triangles. In addition, note that since $\Exp{N_k} > 0$ a case distinction for $N_k$ is no longer needed for $c^\ast(k;G)$. It is however still relevant since we are eventually interested in $c(k;G)$. Following the notation convention, throughout the remainder of this paper we write $c^\ast(k; \GPo)$ and $c^\ast(k; \Gbox)$ to denote the modified local clustering function in $\GPo$ and $G_{\mathcal{P},n}(\alpha,\nu)$, respectively.
%
%Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:mainktoinfty} based on the different propositions described below, plus the sections in which theses propositions are proved. Observe that the order in which the intermediate results are proved is reversed with respect to the natural order of reasoning. This does not create any circular logic, since each intermediate result is independent of the others. We choose this order because results proved in the later stages are helpful to deal with error terms coming up in proofs at earlier stages and hence help streamline those proofs. Below we briefly describe each of the intermediate steps leading up to the proof of Theorem~\ref{thm:mainktoinfty}. We start with an observation on the dependence between the degree of a point $p = (x,y)$ and its height $y$.
%>>>>>>> Stashed changes

@@

\caption{Overview of the proof strategy for Theorem \ref{thm:local_clustering_hyperbolic}. The left column denote the models in which the true hyperbolic balls are used while the right column contains the models that use an approximation of these. The most important part is the transition between these to setting which is accomplished by Proposition~\ref{prop:couling_c_H_P}.}
\label{fig:overview_proof}
\end{figure}

@@

\begin{proof}[Proof of Theorem~\ref{thm:mainktoinfty}]
Note that the second of the theorem follows immediately from the first.

@@

All of this together yields that:
\begin{align*}
    \Exp{\left|c(k_n;G_n)-\gamma(k_n)\right|} = \smallO{s(k_n)} = \smallO{\gamma(k_n)},
\end{align*}
which establishes the first statement of the theorem and finishes the proof.  
\end{proof}

>>>>

Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:mainktoinfty} based on the different propositions described below, plus the sections in which theses propositions are proved. Observe that the order in which the intermediate results are proved is reversed with respect to the natural order of reasoning. This does not create any circular logic, since each intermediate result is independent of the others. We choose this order because results proved in the later stages are helpful to deal with error terms coming up in proofs at earlier stages and hence help streamline those proofs. Below we briefly describe each of the intermediate steps leading up to the proof of Theorem~\ref{thm:mainktoinfty}.

%We start with an observation on the dependence between the degree of a point $p = (x,y)$ and its height $y$.
%=======
%Notice that the only difference between $c(k;G)$ and $c^\ast(k;G)$ is that we replace $N_k$ by its expectation $\Exp{N_k}$. The advantage is that now, the only randomness is in the formation of triangles. In addition, note that since $\Exp{N_k} > 0$ a case distinction for $N_k$ is no longer needed for $c^\ast(k;G)$. It is however still relevant since we are eventually interested in $c(k;G)$. Following the notation convention, throughout the remainder of this paper we write $c^\ast(k; \GPo)$ and $c^\ast(k; \Gbox)$ to denote the modified local clustering function in $\GPo$ and $G_{\mathcal{P},n}(\alpha,\nu)$, respectively.
%
%Figure~\ref{fig:overview_proof} shows a schematic overview of the proof of Theorem~\ref{thm:mainktoinfty} based on the different propositions described below, plus the sections in which theses propositions are proved. Observe that the order in which the intermediate results are proved is reversed with respect to the natural order of reasoning. This does not create any circular logic, since each intermediate result is independent of the others. We choose this order because results proved in the later stages are helpful to deal with error terms coming up in proofs at earlier stages and hence help streamline those proofs. Below we briefly describe each of the intermediate steps leading up to the proof of Theorem~\ref{thm:mainktoinfty}. We start with an observation on the dependence between the degree of a point $p = (x,y)$ and its height $y$.
%>>>>>>> Stashed changes

@@

\caption{Overview of the proof strategy for Theorem \ref{thm:mainktoinfty}. The left column denote the models in which the true hyperbolic balls are used while the right column contains the models that use an approximation of these. The most important part is the transition between these to setting which is accomplished by Proposition~\ref{prop:couling_c_H_P}.}
\label{fig:overview_proof}
\end{figure}

@@

\begin{proofof}{Theorem~\ref{thm:mainktoinfty}}
Note that the second of the theorem follows immediately from the first.

@@

All of this together yields that:
\begin{align*}
    \Exp{\left|c(k_n;G_n)-\gamma(k_n)\right|} = \smallO{s(k_n)} = \smallO{\gamma(k_n)},
\end{align*}
which establishes the first statement of the theorem and finishes the proof.  
\end{proofof}

### 2020-03-11 transition.tex
small corrections

\begin{proof}[Proof of Proposition~\ref{prop:convergence_average_clustering_P_n}]

@@

\end{tikzpicture}
\caption{Example for a given $p$ of the boundary function $x^\prime \mapsto b^\ast_p(x^\prime)$, given by the red curve, which determines whether $\mathcal{T}(p,p^\prime) = \emptyset$. We see that when $y^\prime = b^\ast_p(x^\prime)$ then $(\hat{x}(p,p^\prime), \hat{y}(p,p^\prime)) = (x^\ast(p^\prime), y^\ast(\prime))$.}
\label{fig:comparing_triangles_diff_analysis}
\end{figure}

@@

\begin{proof}[Proof of Lemma \ref{lem:clustering_error_T_term}]
Due to symmetry it is enough to show that
\begin{equation}\label{eq:clustering_error_T_main}
	\int_0^{R}\int_0^{I_n} \mu\left(\mathcal{T}(p,p_1)\right) \dd \mu(p_1) = \bigO{y n^{-(2\alpha - 1)} + n^{-(2\alpha-1)} e^{y}}
\end{equation}
The proof goes in two stages. First we compute $\mu\left(\mathcal{T}(p,p_1)\right)$ by splitting it over three disjoint regimes with respect to $p_1$, with $x_1 \ge 0$. Then we do the integration with respect to $p_1$.

>>>>

\begin{proofof}{Proposition~\ref{prop:convergence_average_clustering_P_n}}

@@

\end{tikzpicture}
\caption{Example for a given $p$ of the boundary function $x^\prime \mapsto b^\ast_p(x^\prime)$, given by the red curve, which determines whether $\mathcal{T}(p,p^\prime) = \emptyset$. We see that when $y^\prime = b^\ast_p(x^\prime)$ then $(\hat{x}(p,p^\prime), \hat{y}(p,p^\prime)) = (x^\ast(p^\prime), y^\ast(p^\prime))$.}
\label{fig:comparing_triangles_diff_analysis}
\end{figure}

@@

\begin{proofof}{Lemma \ref{lem:clustering_error_T_term}}
Due to symmetry it is enough to show that
\begin{equation}\label{eq:clustering_error_T_main}
	\int_0^{R}\int_0^{I_n} \mu\left(\mathcal{T}(p,p_1)\right) \dd \mu(p_1) = \bigO{y n^{-(2\alpha - 1)} + n^{-(2\alpha-1)} e^{y}}
\end{equation}
The proof goes in two stages. First we compute $\mu\left(\mathcal{T}(p,p_1)\right)$ by splitting it over three disjoint regimes with respect to $p_1$, with $x_1 \ge 0$. Then we do the integration with respect to $p_1$.
